---
title: "Data Analysis About Academic Calandar"
author: "2Lazy2BeOriginal"
date: "2024-12-24"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)
```

This document used Python + Selenium to collect the course codes and their respective department and faculty. We are interested in finding the most common course code among all the classes and within each faculty. In addition, this will iterate through each department and compare to the other departments to count how many codes are shared among the two. 

**Note**: There are many courses that are no longer offered at the UofM and are still counted in this list. The Academic Calendar does not differentiate between which current classes offered or not. 

```{r create table and Vec}
# use read.delim for sep that is not space or comma

setwd("/Users/2lazy2beoriginal/PycharmProjects/GetAcademicCalandar")
# /Users/2lazy2beoriginal/PycharmProjects/GetAcademicCalandar/output2.txt
factors <- c("dept", "code", "faculty", "courseName")
all_courses <- read.delim("./output2.txt", sep = "|", col.names = factors)
# rename some faculty to be better
all_courses$faculty[all_courses$faculty  == "management-business"] <- "business"
all_courses$faculty[all_courses$faculty  == "agricultural-food-sciences-agriculture"] <- "agriculture"
all_courses$faculty[all_courses$faculty  == "art"] <- "school-of-art"
all_courses$faculty[all_courses$faculty  == "rady"] <- "health sciences"
all_courses$faculty[all_courses$faculty  == "kinesiology-recreation-management"] <- "kin-rec"

all_faculties <- unique(all_courses$faculty)
all_dept <- unique(all_courses$dept)

numFacs <- length(all_faculties)
```

## What is the most common course number? 

```{r finding top 5}
sortedCodes <- sort(table(code = all_courses$code), decreasing = TRUE)
sortedCodes[1:15]
```

```{r plotTopCodes;groupByFac,  fig.width = 18, fig.height = 7}
library('forcats')
library(ggplot2)
facOrder <- c("arts","engineering", "health sciences", "science", "other")
topCodes <- names(sortedCodes[1:15])
topCodesDf <- all_courses[which(all_courses$code %in% topCodes),]
topCodesDf <- topCodesDf[order(topCodesDf$faculty),]
# prevent R from thinking these are numbers and treat as string
topCodesDf$code <- as.character(topCodesDf$code)
# make colour coding less annoying
cherryPickFac <- c("asper", "engineering", "science", "health sciences", "arts")
# show the plot
topCodesDf$faculty[!topCodesDf$faculty %in% cherryPickFac] <- "other"
# reorder so other is at the bottom
topCodesDf$faculty = factor(topCodesDf$faculty, facOrder)

  ggplot(topCodesDf, aes(code)) + geom_bar(aes(fill = faculty)) +
  labs(title = "Most common course codes",
       subtitle = "colour coded by faculty") + 
   theme_gray(base_size = 22)
```

## Table of the courses and the names 

```{r showCoursesUnderAcode ,results = 'asis', warning=FALSE, echo = FALSE}
# this threw a depreciated with xfun::attr
#topCodesDf <- all_courses[which(all_courses$code %in% top5),]
#  topCodesDf %>% 
#    dplyr::group_by(code) %>% 
#  kbl(row.names = FALSE, align = "c") %>% 
#    add_header_above(c(plotTitle = 4)) %>% 
#      kable_styling() %>% 
#    print()
  
library(knitr)
library(kableExtra)
makeTableForTopCodes <- function(x) {
  all4010 <- all_courses[which(all_courses$code == x),c("dept", "faculty", "courseName")]
  all4010 <- all4010[order(all4010$faculty),]
  all4010 <- cbind(1:nrow(all4010),all4010)
  plotTitle <- paste("Courses with code ",x)
  colnames(all4010) <- c("Count", "Department", "Faculty", "Course Name")
  kbl(all4010, row.names = FALSE, align = "c") %>% 
    add_header_above(c(setNames(4, paste("Department with course code", x, " and name of the course")))) %>% 
      kable_styling() %>% 
    print()
}

top5 <- names(sortedCodes[1:5])
for (i in 1:5) {
  makeTableForTopCodes(top5[i])
}
```
So (perhaps surprisingly) the most common course code is not 1000 but 4010 with 24 departments using it in their classes. 

For many suprised by this (particualrly those who wonder why 1010, 1020 are not more common)

4 science courses have 1010 and 1 econ 

## What is the most common course number by faculty

```{r findMostCommonByFac, echo = FALSE}
library(stringr)
# get the top 5 classes in each faculty, excluding law and music
getMostCommonCodeByFacs <- function(fac) {
  tempDf <- all_courses[which(all_courses$faculty == fac),]
  sort(table(code = tempDf$code), decreasing = TRUE)[1:5]
}
moreOneFacVec <- all_faculties[!all_faculties %in% c("music", "law", "social-work")]
splitByFacCodes <- lapply(moreOneFacVec,getMostCommonCodeByFacs)
names(splitByFacCodes) <- moreOneFacVec
# wrap into a Df 
mostCommonByFacDf <- t(as.data.frame(splitByFacCodes))
# duplicate so we can table the data
moreOneFacVec2 = c()
for (i in 1:length(moreOneFacVec)) {
  moreOneFacVec2 <- c(moreOneFacVec2, moreOneFacVec[i], moreOneFacVec[i])
}
mostCommonByFacDf <- cbind(rep(c("Code", "Frequency"), length(moreOneFacVec)),mostCommonByFacDf,moreOneFacVec2)
```

```{r showTableOfCommonCodes, echo = FALSE, warning=FALSE}
library(kableExtra)
# -7 is the column of the groups
kbl(mostCommonByFacDf[,-7], row.names = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover")) %>% 
# add headers
add_header_above(c(" " = 1, "1st" = 1, "2nd" = 1, "3rd" = 1, "4th" = 1, "5th" = 1)) %>% add_header_above(c("Faculty", "Placements by common frequency" = 5)) %>% 
add_header_above(c("Most common course codes grouped by Department" = 6)) %>%
# add colours to the cols
  column_spec(2, background = "#FFDF00") %>%
  column_spec(3, background = "silver") %>%
  column_spec(4, background = "#CE8946") %>%
# add rowNames
pack_rows(index = table(moreOneFacVec2), label_row_css = "background-color: #666; color: #fff;") %>%
# add footnote omiting
footnote(general = "music, law, and social-work are omitted as those faculties have only have one course prefix")
```
The arts faculty is the largest so it makes sense that it has the most departments with the same course codes. There are 

## Which department has the most classes? 

```{r show dept class count top 60, fig.width = 25}
# Do later but add porportion of courses offered
library(gt)
suppressMessages(library(tidyverse))

sortByDept <- as.data.frame(sort(table(Department = all_courses$dept),decreasing = TRUE)[1:60])
# add porportion and count
sortByDept <- cbind(Rank = 1:60, sortByDept, "Relative Porportion" = round(sortByDept$Freq / nrow(all_courses),digits = 3) )

suppressMessages(a <- bind_cols(sortByDept[1:10,], sortByDept[11:20,],sortByDept[21:30,]))
suppressMessages(b <- bind_cols(sortByDept[31:40,], sortByDept[41:50,],sortByDept[51:60,]))
gt_tbl <- gt(a)
gt_tbl <- 
  gt_tbl |>
    opt_stylize(style = 1) |> 
    tab_spanner("1st-10th", 1:4) |>
    tab_spanner("11th-20th", 5:8) |>
  tab_spanner("21st-30th", 9:12) |>
  tab_header(
    title = md("**The departments with number of courses offered**"),
    subtitle = md("These may include courses *no longer availiable*")
  ) |> 
  cols_label(
    "Rank...1" = "Rank",
    "Department...2" = "Department",
    "Freq...3" = "Freq",
    "Relative Porportion...4" = "Relative Porportion",
    "Rank...5" = "Rank",
    "Department...6" = "Department",
    "Freq...7" = "Freq",
    "Relative Porportion...8" = "Relative Porportion",
    "Rank...9" = "Rank",
    "Department...10" = "Department",
    "Freq...11" = "Freq",
    "Relative Porportion...12" = "Relative Porportion"
  )
# Show the gt Table
gt_tbl
# show it again but directly below
gt_tbl <- gt(b)
gt_tbl <- 
  gt_tbl |>
    opt_stylize(style = 1) |> 
    tab_spanner("31st-40th", 1:4) |>
    tab_spanner("51st-50th", 5:8) |>
  tab_spanner("51st-60th", 9:12) |>
  cols_label(
    "Rank...1" = "Rank",
    "Department...2" = "Department",
    "Freq...3" = "Freq",
    "Relative Porportion...4" = "Relative Porportion",
    "Rank...5" = "Rank",
    "Department...6" = "Department",
    "Freq...7" = "Freq",
    "Relative Porportion...8" = "Relative Porportion",
    "Rank...9" = "Rank",
    "Department...10" = "Department",
    "Freq...11" = "Freq",
    "Relative Porportion...12" = "Relative Porportion"
  )
# Show the gt Table
gt_tbl
```

## Porportion of classes by faculty

```{r pie chart, echo = FALSE, fig.width = 14, fig.height = 14}
# unique(sort) is just luck, cex increase text size but may not appear right on a phone
pie(table(all_courses$faculty), 
    label = paste(unique(sort(all_courses$faculty))," ", round(prop.table(table(all_courses$faculty))*100), "%"), col=rainbow(length(all_faculties)), 
    radius = 0.8, cex=1.5)

```

## What is the shortest and longest course name

```{r find shortest/longest course names}
# do average word length by faculty/dept and also most common words

library(tokenizers)
# Count number of words in each courseName
word_count <- sapply(strsplit(all_courses$courseName, "\\s+"), length)
# Count number of characters in each courseName
char_count <- nchar(all_courses$courseName)

withWordCountDf <- all_courses[order(word_count, char_count), ]
withWordCountDf <- cbind(withWordCountDf, numWords = sort(word_count))

# get top 15 shortest and longest course names
head(withWordCountDf, 15)
tail(withWordCountDf, 15)
# find mean word length of each class 
meanWordLen <- sapply(all_dept, function(x) {
  mean(withWordCountDf[which(withWordCountDf$dept == x), "numWords"])
})
# normal
hist(meanWordLen)
# right skewed, perhaps gamma 
hist(word_count)

library(MASS)
fitdistr(char_count, "gamma", start=list(shape=1, rate=1))$estimate

justArts <- withWordCountDf[which(withWordCountDf$faculty == "arts"),"numWords"]
library(DescTools)
DescTools::PlotQQ(word_count, function(p) qexp(p, 1 / 4.14))
DescTools::PlotQQ(word_count, function(q) qgeom(q, 1 - exp(-1/4.14)))
DescTools::PlotQQ(char_count, function(p) qgamma(p, 5.1823 / 0.169))
DescTools::PlotQQ(justArts, function(p) qpois(p, 4.32))
```
## Does it follow Benfords Law 

## What is the most common word 

```{r findMostCommonWord}
# 1. Collapse all strings into one
all_text <- paste(withWordCountDf$courseName, collapse = " ")
# 2. Split into individual words
words <- unlist(strsplit(all_text, "\\s+"))
# 3. Convert to lowercase (optional, to avoid case sensitivity)
words <- tolower(words)
# 4. Count word frequency
word_freq <- table(words)
sort(word_freq, decreasing = TRUE)
```


## What happens if you discard most of the words 

```{r discard words} 
filter_string_by_words <- function(w, s) {
  # Initialize a data frame to store results
  results <- data.frame(filteredString = character(length(s)),
                        originalString = character(length(s)),
                        wordsRemoved = integer(length(s)),
                        stringsAsFactors = FALSE)
  
  for (i in seq_along(s)) {
    # Split the string into words
    wd <- unlist(strsplit(s[i], "\\s+"))
    
    # Keep only words in w
    kept_words <- wd[wd %in% w]
    
    # Reconstruct filtered string
    filtered_str <- paste(kept_words, collapse = " ")
    
    # Count removed words
    words_removed <- length(wd) - length(kept_words)
    
    # Store in results data frame
    results$filteredString[i] <- filtered_str
    results$wordsRemoved[i] <- words_removed
    results$originalString[i] <- s[i]
  }
  
  return(results)
}

# get all the words that show up at least 5 times 
topWordFreq <- names(word_freq[which(word_freq > 1)])
res <- filter_string_by_words(topWordFreq, tolower(all_courses$courseName))
```
## 2890, 3890, 4890 are co op work terms 

## Which department is most similar to one another, how about faculty

* use document distance 

## Which word is the most infomation dense

* some words are common but don't narrow down by deoartment (such as and and (1))
* some words are not common but heavily narrow down department but have to get "lucky" (like statistics narrowing down to just STAT and phys but that's it)

* Use the infomation theory from 3B1B to narrow it down 


## Which department shares the most classes with another


the `table()` function in R groups repeated entries and returns the frequency. Each department has a unique code that cooresponds to a particular class. (MATH 1500 always means Introduction to Calculus, PHIL 1300 always means Introduction to Logic). In other words, each department course code only appears once.

And if we combine the two departments and table the data, a frequency of two automatically implies the code exists in both departments. Then we simply count how many codes have a count of two.

```{r createMostCommonMatrix, echo = FALSE}
# BIOE appears twice in both agriculture and bio engineering

# Only departments with ten or more classes are considered 
largeDeptVec <- names(which(table(all_courses$dept) > 10))
# symmetric matrix where each i,j entry is a pair of departments
mostCommonMatrix = matrix(nrow = length(largeDeptVec), ncol = length(largeDeptVec), dimnames = list(largeDeptVec, largeDeptVec))

# iterate through each possible pair
for (y in 1:length(largeDeptVec)) {
  for (x in 1:length(largeDeptVec)) {
    # name for department 1 and 2 
    target1 <- colnames(mostCommonMatrix)[[x]]
    target2 <- colnames(mostCommonMatrix)[[y]]
    # combine the two department codes
    combinedVec = c(all_courses$code[all_courses$dept == target1], all_courses$code[all_courses$dept == target2])
    temp <- table(combinedVec)
    # The case where x = y is same course counts the number of courses in a department
    mostCommonMatrix[y,x] <- sum(temp == 2)
  }
}

commonMatTest = mostCommonMatrix
diag(commonMatTest) = rep(0, ncol(mostCommonMatrix))
                          
which(commonMatTest == max(commonMatTest), arr.ind=TRUE)
max(commonMatTest)
```

The most two departments with the most course codes in common among the two are **History** and **Religion** with 26 codes shared among the two

## What about the other departments?

```{r calc other dept similarirty}
# df with dept name, matching dept and count of same course codes
indices = data.frame(from = largeDeptVec, to = NA, freq = NA)

# go through the columns and take the max
for (i in 1:length(largeDeptVec)) {
  # each row is a department
  currRow <- commonMatTest[i,]
  # we order the data since it tells us the index since it is how we backtrack to the department name
  keyIndex <- order(currRow, decreasing = TRUE)[1]
  matchingDept <- colnames(commonMatTest)[keyIndex]
  count <- max(currRow)
  
  # add to the list 
  indices$to[i] <- matchingDept
  indices$freq[i] <- count
}

# this df is sorted by frequency most to lowest. We will skip every other entry since it will be the mirrored version

indices <- indices[order(indices$freq, decreasing = TRUE),]
indices <- indices[seq(1, nrow(indices), by = 2),]
```

```{r echo = FALSE}
library(gt)
gt_tbl <- gt(head(indices, 25))
gt_tbl <- 
  gt_tbl |>
  #dplyr::filter(Freq >= 10) 
  tab_header(
    title = md("**Each department and a different department corresponding most common course code**"),
  )
# Show the gt Table
gt_tbl
```
